Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1]}

### 2018-01-23 paper/paper.tex
Little bits

\begin{abstract}
    Modern subgraph isomorphism solvers use degree-based heuristics to direct a constraint
    programming style search. We show that discrepancy-based searches are too expensive to be
    effective in offsetting mistakes made by value-ordering heuristics in this situation.  We then
    demonstrate that a random heuristic with restarts beats conventional search with the best-known
    tailored heuristic. This motivates the introduction of a slightly-random heuristic with similar
    behaviour to the tailored heuristic: when combined with restarts, this new search strategy is
    over a hundred times more effective on satisfiable instances, and does not affect performance on
    unsatisfiable instances. Finally, we show that the same strategy also improves two maximum
    common induced subgraph algorithms.
\end{abstract}

@@

These large overheads are to be expected: discrepancy searches are aimed primarily at getting better
feasible solutions in optimisation problems which are too large for a proof of optimality to be a
realistic prospect, and they are not well-suited for unsatisfiable decision problems. However, the
extremely large gains on some satisfiable instances confirm our suspicions that we should look at
value-ordering heuristics in more detail.

@@

@@

In \cref{figure:value-ordering-heuristics} we show the cumulative number of satisfiable instances
solved over time using four different value-ordering heuristics:

\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
current state of the art \citep{DBLP:conf/ijcai/McCreeshPT16}.

@@

In the more detailed view in \cref{figure:scatter-by-family}, comparing the basic algorithm with the
degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
other hand, there are ?? large numbers of satisfiable instances well below the diagonal line,
indicating large speedups.  Better yet, there are only ?? satisfiable instances that are more than a
factor of ?? times worse.

?? By family

?? This has made up the consistency we lost from introducing randomness.

?? Explain why this is so amazingly awesome and the most important idea ever to happen in the entire
history of science.

?? Justify aggregate speedups on sat instances

@@

?? Two recent maximum common induced subgraph algorithms: k${\downarrow}$
\citep{DBLP:conf/aaai/HoffmannMR17}, and McSplit \citep{DBLP:conf/ijcai/McCreeshPT17}. The former
attempts to solve the problem by first trying to solve the induced subgraph isomorphism problem, and
then if that fails, retries allowing a single unmatched vertex (and thus using weaker invariants),
and so on. Due to its similarity to \cref{algorithm:sip}, we can introduce the same bias and restart
strategy, requiring only a small amount of jiggery-pokery to handle wildcards in nogoods.

The McSplit algorithm, meanwhile, uses a constraint programming style search, but with special
propagators and backtrackable data structures that exploit special properties of the problem. The
unconventional domain store used by McSplit precludes the use of arbitrary unit propagation, and so
when introducing restarts, we cannot propagate using nogoods. Instead, we can only detect when we
are inside an already-visited branch. We must therefore use the one watched literal scheme instead.

@@

\clearpage
\appendix

\section{Extra Plots not for the Final Paper}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-as.pdf}
    \caption{?? Aggregate speedups on satisfiable instances}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-as-mcs.pdf}
    \caption{?? Aggregate speedups on MCS instances}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-degree.pdf}

    \caption{When is the degree heuristic good?}
    \label{figure:scatter-degree}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-constant.pdf}

    \caption{Constant restarts?}
    \label{figure:scatter-constant}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-parallel.pdf}

    \caption{Parallel}
    \label{figure:parallel}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-parallel.pdf}

    \caption{Parallel Luby}
    \label{figure:parallel-luby}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-parallel-constant.pdf}

    \caption{Parallel Constant}
    \label{figure:parallel-constant}
\end{figure}

>>>>

\begin{abstract}
    Modern subgraph isomorphism solvers use degree-based value-ordering heuristics to direct
    backtracking search. We show that discrepancy-based searches are too expensive to be effective
    in offsetting heuristic mistakes in this situation.  We then demonstrate that a random heuristic
    with restarts beats conventional search with the best-known tailored heuristic. This motivates
    the introduction of a slightly-random heuristic with similar behaviour to the tailored
    heuristic: when combined with restarts, this new search strategy is over a hundred times more
    effective on satisfiable instances, and does not affect performance on unsatisfiable instances.
    Finally, we show that this strategy also improves two maximum common induced subgraph
    algorithms.
\end{abstract}

@@

@@

\Cref{figure:value-ordering-heuristics} shows the cumulative number of satisfiable instances
solved over time using different value-ordering heuristics:

\paragraph{Degree} We select vertices from highest degree to lowest degree
\citep{DBLP:conf/ijcai/McCreeshPT16}.

@@

These large overheads are to be expected: discrepancy searches are aimed primarily at getting better
feasible solutions in optimisation problems which are too large for a proof of optimality to be a
realistic prospect, and they are not well-suited for unsatisfiable decision problems. However, the
extremely large gains on some satisfiable instances confirm our suspicions that we should find an
alternative to heavy-commitment backtracking search.

@@

In the more detailed view in \cref{figure:scatter-by-family}, comparing the basic algorithm with the
degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
other hand, there are ?? large numbers of satisfiable instances well below the diagonal line,
indicating large speedups.  Better yet, there are only ?? satisfiable instances that are more than a
factor of ten times worse. In other words, as well as improving performance, we have made up most of
the consistency we lost by introducing randomness.

@@

Two recent maximum common induced subgraph algorithms also make use of backtracking search with
degree as a value-ordering heuristic. The k${\downarrow}$ algorithm
\citep{DBLP:conf/aaai/HoffmannMR17} attempts to solve the problem by first trying to solve the
induced subgraph isomorphism problem, and then if that fails, retries allowing a single unmatched
vertex (and thus using weaker invariants), and so on. Due to its similarity to \cref{algorithm:sip},
we can introduce the same bias and restart strategy, requiring only a small amount of jiggery-pokery
to handle wildcards in nogoods.  Meanwhile, the McSplit algorithm
\citep{DBLP:conf/ijcai/McCreeshPT17} uses a constraint programming style search, but with special
propagators and backtrackable data structures that exploit special properties of the problem. The
unconventional domain store used by McSplit precludes the use of arbitrary unit propagation, and so
when introducing restarts, we cannot propagate using nogoods.  Instead, we can only detect when we
are inside an already-visited branch.  We must therefore use the one watched literal scheme instead.

@@
